---
title: "FUS regulates DNA replication timing in U-2 OS cells"
author: "Weiyan"
date: "4/2/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/weiyanjia/Desktop/FUS_paper_Figures/FUS_Paper")
```

**Note**:

> Asynchronous U-2 OS cells and three FUS deletion clones (Clone46, Clone65 and Clone110) were pulse-labeled with EdU and then the S phase cells were divided to three stages based on EdU foci pattern as shown in U-2 OS. 

```{r}
getwd()
```
# 1.Barplot of RT in FUS knock out cells
## 1.1 data organization

```{r}
RT_1 <- read.delim("clean/RT_1.txt", header = TRUE, sep ="\t", stringsAsFactors = FALSE)
RT_1
library(tidyr)
RT_1_sum <- RT_1 %>%
  gather(key = sample, value = cell_count, -RT)
RT_1_sum
library(dplyr)
RT_1_sum <- RT_1_sum %>%
  group_by(sample, RT) %>%
  summarise(cell_count= sum(cell_count))
RT_1_sum
library(data.table)
setDT(RT_1_sum)[, Percentage := round(100*cell_count/sum(cell_count), 2),
                   by =  sample]
str(RT_1_sum)
RT_1_sum$RT <- factor(RT_1_sum$RT, levels= c("Late S","Middle S","Early S"))
RT_1_sum$sample <- factor(RT_1_sum$sample, levels= c("Clone110","Clone65","Clone46","U2OS"))

RT_1_sum
```

## 1.2 barplot of RT_1

```{r RT of FUS Knockout cells and reconstitued cells, fig.height=4, fig.width=6}
library(dplyr)
library(tidyverse)
RT_2 <- read.delim("clean/RT_2.txt", header = TRUE, sep ="\t", stringsAsFactors = FALSE)
RT_2
RT_2_sum <- RT_2 %>%
  gather(key = sample, value = cell_count, -RT) # for simple : gather(sample, cell_count, -RT)

RT_2_sum

RT_2_sum <- RT_2_sum %>%
  group_by(sample) %>%
  mutate(Percentage= round(100*cell_count/sum(cell_count),2))
RT_2_sum 

RT_2_sum$RT <- factor(RT_2_sum$RT, levels= c("Late S","Middle S","Early S"))
RT_2_sum$sample <- factor(RT_2_sum$sample, levels= c("FUSClone110","GUSClone110","FUSU2OS","GUSU2OS"))
RT_2_sum
write.csv(RT_2_sum, file="clean/RT_EdU_FUS-KO.csv")
library(ggpubr)
ggbarplot(RT_2_sum, x="sample", 
          y= "Percentage",
          ylab = "Percentage(%)",
          xlab = "",
          legend.title ="",
          color = "RT",
          fill = "RT",
          palette = "aaas",
          label = TRUE,
          lab.pos = "in",
          width = 0.6,
          orientation = "horiz",
          lab.col = "white"
          )+   
  scale_y_continuous(expand = c(0, 0))
```

# 2. Barplot of RT using Double thymidine treated samples

## 2.1 data loading
note: 1. Asychronized sample named by "0" in time column.
      2. time was relase from doublt thymidine treatement (hours)

```{r}
RT_dt <- read.delim("clean/RT_DT_assay.txt", header = TRUE, sep ="\t", stringsAsFactors = FALSE)
RT_dt
RT_dt <- RT_dt %>%
  group_by(Sample, time) %>%
  mutate(Percentage= round(100*cell_count/sum(cell_count),2))
RT_dt 
RT_dt$RT <- factor(RT_dt$RT, levels= c("Late S","Middle S","Early S"))
RT_dt$sample_order <- factor(RT_dt$Sample, levels= c("FUSClone110","GUSClone110","GUSU2OS"))
RT_dt
write.csv(RT_dt, file="clean/RT_BrdU_DoubleThymidine_FUS.csv")
```
## 2.2 barplot

### 2.2.1 barplot of Double thymidine data

```{r}
library(ggpubr)
dt<- ggbarplot(RT_dt, x="sample_order", 
          y= "Percentage",
          facet.by = "time",
          ylab = "Percentage(%)",
          xlab = "",
          legend.title ="",
          color = "RT",
          fill = "RT",
          palette = "aaas",
          label = TRUE,
          lab.pos = "in",
          width = 0.6,
          orientation = "horiz",
          lab.col = "white"
          )+   
  scale_y_continuous(expand = c(0, 0))
dt
```

### 2.2.2 barplot of Double thymidine data (facet by time)
```{r fig.height=8, fig.width=6}
facet(dt, facet.by="time", ncol= 1,
      panel.labs.background = list(fill = "white", color = "grey")
     )
```
