---
title: "Quantification results of PLA_FUS"
author: "Weiyan"
date: "4/2/2020"
output: github_document
---

>Check the interaction between FUS, and POLD1,FEN1 and PCNA under with preextraction

**Note**:

> 1. Samples: GUS/U2OS(FUS, WT) and GUS/Clone110(FUS,KO);
  2. Preextraction with CKS+0.5% Triton-X100 8min on ice with shacking (speed: 4)
  3. Antibodies: FUS, POLD1,FEN1, PCNA;
  4. PLA foci was quantified by CellProfiler
  5. Data collected on 20190828

# Packages

```{r message=FALSE, warning=FALSE}
library(ggpubr)
library(ggsci)
library(ggExtra)
library(export)
library(tidyverse)
```

# 1. FEN1

## 1.1 **FEN1_data loading**

```{r}
FEN1_WT_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/FEN1_U2OS_CSK_Nuclei.csv",header = TRUE)
FEN1_KO_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/FEN1_Clone110_CSK_Nuclei.csv",header = TRUE)
```

## 1.2 **FEN1_U2OS_CSK**

```{r fig.height=6, fig.width=8}
FEN1_WT_CSK_Nuclei_1<- FEN1_WT_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="WT_CSK")%>%
                      dplyr::mutate(antibody="FEN1")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 1.3 **FEN1_Clone110_KO**

```{r fig.height=6, fig.width=8}
FEN1_KO_CSK_Nuclei_1<- FEN1_KO_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="KO_CSK")%>%
                      dplyr::mutate(antibody="FEN1")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 1.4 **FEN1 Merged data**

```{r}

PLA_FEN1_Nuclei<- bind_rows(
                            FEN1_WT_CSK_Nuclei_1,
                            FEN1_KO_CSK_Nuclei_1
                            )
PLA_FEN1_Nuclei <- PLA_FEN1_Nuclei%>%
                   dplyr::mutate(condition="Pre-extraction")
```

```{r}
PLA_FEN1_Nuclei_Median <- PLA_FEN1_Nuclei%>%
              group_by(sample,antibody,condition)%>%
              summarise(MD = median(Children_PLA_Count))
PLA_FEN1_Nuclei_Median
PLA_FEN1_Tally <-PLA_FEN1_Nuclei%>%
                 group_by(sample)%>%
                 tally()
PLA_FEN1_Tally
```


```{r fig.height=6, fig.width=8}
FEN1_CSK <- PLA_FEN1_Nuclei%>%
             filter(condition == "Pre-extraction")
ggboxplot(FEN1_CSK, 
          x="sample",
          y="Children_PLA_Count",
          color = "sample",
          palette = "aaas",
          add = "median",
          ylab = "PLA Counts",
          xlab = ""
          )+
  stat_compare_means(comparisons = list(c("WT_CSK","KO_CSK")), method = "wilcox.test",
                     label = "p.format",
                     label.y =30)+ ## method:"wilcox.test", "anova"...
  geom_text(data = subset(PLA_FEN1_Nuclei_Median, condition == "Pre-extraction" ), aes(x=sample, y= MD, label = MD),
              position = position_dodge(width = 0.8), size = 3, vjust = -0.5,
            ) 
graph2pdf(file="/Users/weiyanjia/Desktop/PLA/08282019/analysis/results/PLAofFEN1_CSK.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

# 2. PCNA

## 2.1 **PCNA_data loading**

```{r}
PCNA_WT_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/PCNA_U2OS_CSK_Nuclei.csv",header = TRUE)
PCNA_KO_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/PCNA_Clone110_CSK_Nuclei.csv",header = TRUE)
```

## 2.2 **PCNA_U2OS_CSK**

```{r fig.height=6, fig.width=8}
PCNA_WT_CSK_Nuclei_1<- PCNA_WT_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="WT_CSK")%>%
                      dplyr::mutate(antibody="PCNA")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 2.3 **PCNA_Clone110_KO**

```{r fig.height=6, fig.width=8}
PCNA_KO_CSK_Nuclei_1<- PCNA_KO_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="KO_CSK")%>%
                      dplyr::mutate(antibody="PCNA")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 2.4 **PCNA Merged data**

```{r}
PLA_PCNA_Nuclei<- bind_rows(
                            PCNA_WT_CSK_Nuclei_1,
                            PCNA_KO_CSK_Nuclei_1
                            )
PLA_PCNA_Nuclei <- PLA_PCNA_Nuclei%>%
                   dplyr::mutate(condition="Pre-extraction")
```

```{r}
PLA_PCNA_Nuclei_Median <- PLA_PCNA_Nuclei%>%
              group_by(sample,antibody,condition)%>%
              summarise(MD = median(Children_PLA_Count))
PLA_PCNA_Nuclei_Median
PLA_PCNA_Tally <-PLA_PCNA_Nuclei%>%
                 group_by(sample)%>%
                 tally()
PLA_PCNA_Tally
```


```{r fig.height=6, fig.width=8}
PCNA_CSK <- PLA_PCNA_Nuclei%>%
             filter(condition == "Pre-extraction")
ggboxplot(PCNA_CSK, 
          x="sample",
          y="Children_PLA_Count",
          color = "sample",
          palette = "aaas",
          add = "median",
          ylab = "PLA Counts",
          xlab = ""
          )+
  stat_compare_means(comparisons = list(c("WT_CSK","KO_CSK")), method = "wilcox.test",
                     label = "p.format",
                     label.y =80)+ ## method:"wilcox.test", "anova"...
  geom_text(data = subset(PLA_PCNA_Nuclei_Median, condition == "Pre-extraction" ), aes(x=sample, y= MD, label = MD),
              position = position_dodge(width = 0.8), size = 3, vjust = -0.5,
            ) 
graph2pdf(file="/Users/weiyanjia/Desktop/PLA/08282019/analysis/results/PLAofPCNA_CSK.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

# 3. POLD1

## 3.1 **POLD1_data loading**

```{r}
POLD1_WT_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/POLD1_U2OS_CSK_Nuclei.csv",header = TRUE)
POLD1_KO_CSK_Nuclei <- read.csv("/Users/weiyanjia/Desktop/PLA/08282019/analysis/raw/POLD1_Clone110_CSK_Nuclei.csv",header = TRUE)
```

## 3.2 **POLD1_U2OS_CSK**

```{r fig.height=6, fig.width=8}
POLD1_WT_CSK_Nuclei_1<- POLD1_WT_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="WT_CSK")%>%
                      dplyr::mutate(antibody="POLD1")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 3.3 **POLD1_Clone110_KO**

```{r fig.height=6, fig.width=8}
POLD1_KO_CSK_Nuclei_1<- POLD1_KO_CSK_Nuclei%>%
                      dplyr::select(ImageNumber, Children_PLA_Count,Intensity_IntegratedIntensity_DAPI)%>%
                      dplyr::mutate(ImageNumber= as.factor(ImageNumber))%>%
                      dplyr::mutate(sample="KO_CSK")%>%
                      dplyr::mutate(antibody="POLD1")%>%
                      dplyr::mutate(sample= as.factor(sample),antibody= as.factor(antibody))
```

## 3.4 **POLD1 Merged data**

```{r}
PLA_POLD1_Nuclei<- bind_rows(
                            POLD1_WT_CSK_Nuclei_1,
                            POLD1_KO_CSK_Nuclei_1
                            )
PLA_POLD1_Nuclei <- PLA_POLD1_Nuclei%>%
                   dplyr::mutate(condition="Pre-extraction")
```

```{r}
PLA_POLD1_Nuclei_Median <- PLA_POLD1_Nuclei%>%
              group_by(sample,antibody,condition)%>%
              summarise(MD = median(Children_PLA_Count))
PLA_POLD1_Nuclei_Median
PLA_POLD1_Tally <-PLA_POLD1_Nuclei%>%
                 group_by(sample)%>%
                 tally()
PLA_POLD1_Tally
```


```{r fig.height=6, fig.width=8}
POLD1_CSK <- PLA_POLD1_Nuclei%>%
             filter(condition == "Pre-extraction")
ggboxplot(POLD1_CSK, 
          x="sample",
          y="Children_PLA_Count",
          color = "sample",
          palette = "aaas",
          add = "median",
          ylab = "PLA Counts",
          xlab = ""
          )+
  stat_compare_means(comparisons = list(c("WT_CSK","KO_CSK")), method = "wilcox.test",
                     label = "p.format",
                     label.y =20)+ ## method:"wilcox.test", "anova"...
  geom_text(data = subset(PLA_POLD1_Nuclei_Median, condition == "Pre-extraction" ), aes(x=sample, y= MD, label = MD),
              position = position_dodge(width = 0.8), size = 3, vjust = -0.5,
            ) 
graph2pdf(file="/Users/weiyanjia/Desktop/PLA/08282019/analysis/results/PLAofPOLD1_CSK.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```


```{r}
sessionInfo()
```