---
title: "GSEA"
author: "Weiyan"
date: "4/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data preparation
**Note:**

> GSEA preranking metric:

> 1. Select top 12000 genes based on adjusted P value
  2. Run genes based on logfc shrink values from DESeq2

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(org.Hs.eg.db)
library(DT)
library(fgsea)
library(data.table)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
library(dendextend)
library(export)
library(ggpubr)
```


```{r}
WT_KO_res <-read.csv("FUS_DESeq2/results/res_WTvsKO_shrunken_tb_entrez.csv",header = TRUE)
normalized_counts<- read.csv("FUS_DESeq2/results/normalized_counts.csv",header = TRUE)
ann_colors = list(
    group = c(WT = "#7570B3", KO = "#E7298A", RE = "#66A61E")
)
ann.sample<-  read.csv("FUS_DESeq2/results/ann.sample.csv") %>%
              column_to_rownames(var="X")
ann.sample_meta<- ann.sample%>%
                  rownames_to_column(var="sampleName")

```

Preranking metric

```{r}
summary(WT_KO_res)
```

Map Ensembl gene IDs to symbol. First create a mapping table
```{r}
ens2symbol <- AnnotationDbi::select(org.Hs.eg.db,
                                    key=as.character(WT_KO_res$gene),
                                    columns="SYMBOL",
                                    keytype="ENSEMBL")
ens2symbol <- as_tibble(ens2symbol)
head(ens2symbol)
```

Now join them
```{r}
WT_KO_res <- left_join(WT_KO_res, ens2symbol, by=c("gene"="ENSEMBL"))%>%
           distinct()%>%
           drop_na()
```

```{r}
WT_KO_X <- WT_KO_res%>%
           drop_na() %>% #whole table wide
           distinct()%>%
           arrange(padj)%>%
           top_n(-12000,padj)
          
# summary(WT_KO_X) 
# WT_KO_X%>%
# group_by(SYMBOL) %>% dplyr::filter(n()>1) %>% summarize(n=n())
```

```{r}
# stat related to Wald statistic and stat value corelected to log2FC, so we also can use the log2FC as input
res2 <- WT_KO_X %>% 
  dplyr::select(SYMBOL, log2FoldChange) %>% 
  drop_na() %>% 
  distinct() %>% 
  group_by(SYMBOL) %>% 
  summarize(log2FoldChange=mean(log2FoldChange))
head(res2)
```

# 2. GSEA by fgsea package

```{r}
ranks <- deframe(res2)
head(ranks, 20)
barplot(sort(ranks, decreasing = T))
```
## 2.1 GO(BP) analysis

```{r}
set.seed(12)
pathways.GO.BP <- gmtPathways("MSigDB/c5.bp.v6.2.symbols.gmt")
WT_KO_GSEA_BP<-fgsea(pathways=gmtPathways("MSigDB/c5.bp.v6.2.symbols.gmt"), ranks, 
                           nperm=100000,
                           minSize = 15,
                           maxSize = 500
                              )%>% 
                            arrange(padj)
```

```{r fig.height=8, fig.width=8}
Go_BP_05<- WT_KO_GSEA_BP %>%
              dplyr::filter(padj <0.01)%>%
              mutate(nLeadingEdge=lengths(leadingEdge),GeneRatio=lengths(leadingEdge)/size, term= substring(pathway,4))%>%
              dplyr::select(-ES,-nMoreExtreme)%>%
              distinct()

fwrite(Go_BP_05, file="results/Go_BP_001.txt", sep="\t", sep2=c("", " ", ""))
              
### NES: positive, upregulated in KO.
p<- ggplot(Go_BP_05, aes(reorder(term, -NES), -NES)) +  ## order pathway based on NES value 
  geom_col(aes(fill=GeneRatio)) +
  # geom_bar(stat="identity",fill = "red")+
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="GO_BP pathways NES from GSEA") + 
  theme_minimal()
# p + scale_fill_gradient(low="goldenrod2", high="firebrick3") ##scale_color_gradient for shape or scale_color_gradientn(colours = rainbow(5))
p + scale_fill_gradient( low= brewer.pal(6, "YlOrRd"),high="firebrick3")
graph2pdf(file="figures/FUS_GSEA_BP.pdf", width=10, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

## 2.2 Heatmap of DNA_REPLICATION pathway leading genes

```{r}
plotEnrichment(pathways.GO.BP[["GO_DNA_REPLICATION"]],
               ranks) + labs(title="GO_DNA_REPLICATION")
# graph2ppt(file="figures/DNA_REPLICATION.Enrichment.pptx", width =8,aspectr=sqrt(2),bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPLICATION.Enrichment.pptx", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

```{r}
DNA_REPLICATION.Leading<-Go_BP_05 %>% 
             dplyr::filter(pathway == "GO_DNA_REPLICATION")%>%
             dplyr::select(leadingEdge)%>%
             unnest()%>% 
             dplyr::rename(symbol=leadingEdge)%>%
             inner_join(normalized_counts, by="symbol")%>% 
             distinct()
DNA_REPLICATION.Leading.all<-DNA_REPLICATION.Leading%>%
                         column_to_rownames(var ="symbol")
### Filter genes based on padj<0.05
DNA_REPLICATION.Leading.sig <-DNA_REPLICATION.Leading %>%
                          inner_join(WT_KO_res, "symbol"="SYMBOL")%>% 
                          dplyr::select(-SYMBOL,-gene,-baseMean,-lfcSE)%>%
                          distinct()%>%
                          dplyr::filter(padj< 0.05)%>%
                         column_to_rownames(var ="symbol")
```

```{r fig.height=8, fig.width=5}
# method for cluster:
# the agglomeration method to be used. This should be (an unambiguous abbreviation of) one of "ward.D", "ward.D2", "single", "complete", "average" (= UPGMA), "mcquitty" (= WPGMA), "median" (= WPGMC) or "centroid" (= UPGMC).

heat_colors <- brewer.pal(6, "YlOrRd")

dna.rep.out.all<- pheatmap(DNA_REPLICATION.Leading.all[2:10], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation = ann.sample, 
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         cutree_rows = 3,
         cutree_cols = 3,
         fontsize_row = 6, 
         height = 20)

dna.rep_col.all<- sort(cutree(dna.rep.out.all$tree_row, k=3))%>%
                 as.factor()%>%
                 data.frame()%>%
                 dplyr::rename(cluster=".")

pheatmap(DNA_REPLICATION.Leading.all[2:10], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation_col = ann.sample, 
         annotation_row = dna.rep_col.all,
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         cutree_rows = 3,
         cutree_cols = 3,
         fontsize_row = 6, 
         height = 20)
# graph2ppt(file="figures/DNA_REPLICATION.Leading.all.pptx", width =8,aspectr=0.5,bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPLICATION.Leading.all.pdf", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")

## filted by padj
dna.rep.out.sig = pheatmap(DNA_REPLICATION.Leading.sig[1:9], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation_col = ann.sample, 
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         cutree_rows = 3,
         cutree_cols = 3,
         fontsize_row = 10, 
         height = 20)
# If you want something like gene-to-cluster assignment, you can 'cut' your row dendrogram into a pre-selected number of groups as follows:
dna.rep_col.sig<- sort(cutree(dna.rep.out.sig$tree_row, k=3))%>%
                 as.factor()%>%
                 data.frame()%>%
                 dplyr::rename(cluster=".")

pheatmap(DNA_REPLICATION.Leading.sig[1:9], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation_col = ann.sample, 
         annotation_row = dna.rep_col.sig,
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         # # kmeans_k = 4,
         # km=3,
         cutree_rows = 3,
         cutree_cols = 3,
         fontsize_row = 10, 
         height = 20)
# graph2ppt(file="figures/DNA_REPLICATION.Leading.sig.pptx", width =8,aspectr=0.5,bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPLICATION.Leading.sig.pdf", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```
## 2.3 DNA_REPLICATION related genes expression

```{r}
dna.rep_col.sig_reads<- dna.rep_col.sig%>%
                        rownames_to_column(var = "symbol")%>%
                        left_join(normalized_counts, by="symbol")%>% 
                        distinct()%>%
                        dplyr::filter(cluster!=1)

gathered.dna.rep_col.sig_reads<-dna.rep_col.sig_reads%>%
                               gather(colnames(dna.rep_col.sig_reads)[4:12], key = "sampleName", value = "normalized_counts") %>%
                               mutate(sampleName=as.factor(sampleName))%>%
                               left_join(ann.sample_meta, by="sampleName")


gathered.dna.rep_col.sig_reads$group<- factor(gathered.dna.rep_col.sig_reads$group, levels=c("WT","KO","RE")) ##change levles for plot
```

```{r}
ggboxplot(gathered.dna.rep_col.sig_reads, x = "symbol",
          y = "normalized_counts",
          title="DNA Replication Related Genes",
          color = "group",
          add = "jitter",
          # shape="sampleName",
          # group.by= "cluster",
          merge = TRUE,
          ylab = "Normalized Counts",
          xlab = "",
          legend= "right",
          x.text.angle = 45,
          palette = "aaas")+
  yscale("log10", .format = FALSE)
# graph2ppt(file="figures/DNA_Replication.Counts.pptx", width =8,aspectr=sqrt(2),font = "Arial",bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_Replication.Counts.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")

```


## 2.4 Heatmap of DNA_REPAIR pathway leading genes

```{r}
plotEnrichment(pathways.GO.BP[["GO_DNA_REPAIR"]],
               ranks) + labs(title="GO_DNA_REPAIR")
# graph2ppt(file="figures/DNA_REPAIR.Enrichment.pptx", width =8,aspectr=sqrt(2),bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPAIR.Enrichment.pdf", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

```{r}
DNA_REPAIR.Leading<-Go_BP_05 %>% 
             dplyr::filter(pathway == "GO_DNA_REPAIR")%>%
             dplyr::select(leadingEdge)%>%
             unnest()%>% 
             dplyr::rename(symbol=leadingEdge)%>%
             inner_join(normalized_counts, by="symbol")%>% 
             distinct()
DNA_REPAIR.Leading.all<-DNA_REPAIR.Leading%>%
                         column_to_rownames(var ="symbol")
### Filter genes based on padj<0.05
DNA_REPAIR.Leading.sig <-DNA_REPAIR.Leading %>%
                          inner_join(WT_KO_res, "symbol"="SYMBOL")%>% 
                          dplyr::select(-SYMBOL,-gene,-baseMean,-lfcSE)%>%
                          distinct()%>%
                          dplyr::filter(padj< 0.05)%>%
                         column_to_rownames(var ="symbol")
```

```{r fig.height=10, fig.width=5}
heat_colors <- brewer.pal(6, "YlOrRd")

dna.repair.out.all<- pheatmap(DNA_REPAIR.Leading.all[2:10], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation = ann.sample, 
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         # # kmeans_k = 4,
         cutree_rows = 3,
         cutree_cols = 3,
         fontsize_row = 6, 
         height = 20)

dna.repair_col.all<- sort(cutree(dna.repair.out.all$tree_row, k=4))%>%
                 as.factor()%>%
                 data.frame()%>%
                 dplyr::rename(cluster=".")

pheatmap(DNA_REPAIR.Leading.all[2:10], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation_col = ann.sample, 
         annotation_row = dna.repair_col.all,
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         # # kmeans_k = 4,
         cutree_rows = 4,
         cutree_cols = 3,
         fontsize_row = 6, 
         height = 20)

# graph2ppt(file="figures/DNA_REPAIR.Leading.sig.pptx", width =8,aspectr=0.5,bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPAIR.Leading.sig.pdf", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

```{r fig.height=8, fig.width=6}
dna.repair.out.sig<- pheatmap(DNA_REPAIR.Leading.sig[1:9], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation = ann.sample, 
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row",
         # # kmeans_k = 4,
         cutree_rows = 2,
         cutree_cols = 3,
         fontsize_row = 10, 
         height = 20)

dna.repair_col.sig<- sort(cutree(dna.repair.out.sig$tree_row, k=2))%>%
                 as.factor()%>%
                 data.frame()%>%
                 dplyr::rename(cluster=".")

pheatmap(DNA_REPAIR.Leading.sig[1:9], 
         # color = heat_colors,
         clustering_method = "ward.D2",
         cluster_rows = T,
         cluster_cols = T,
         show_rownames =T ,
         annotation_col = ann.sample, 
         annotation_row = dna.repair_col.sig,
         annotation_colors = ann_colors,
         border_color = NA, 
         fontsize = 10, 
         scale = "row",
         # # kmeans_k = 4,
         cutree_rows = 2,
         cutree_cols = 3,
         fontsize_row = 10, 
         height = 20)
# graph2ppt(file="figures/DNA_REPAIR.Leading.sig.pptx", width =8,aspectr=0.5,bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPAIR.Leading.sig.pdf", width=5, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

## 2.5 DNA_REPAIR related genes expression

```{r}
dna.repair_col.sig_reads<- dna.repair_col.sig%>%
                        rownames_to_column(var = "symbol")%>%
                        left_join(normalized_counts, by="symbol")%>% 
                        distinct()%>%
                        dplyr::filter(cluster!=1)

gathered.dna.repair_col.sig_reads<-dna.repair_col.sig_reads%>%
                               gather(colnames(dna.rep_col.sig_reads)[4:12], key = "sampleName", value = "normalized_counts") %>%
                               mutate(sampleName=as.factor(sampleName))%>%
                               left_join(ann.sample_meta, by="sampleName")


gathered.dna.repair_col.sig_reads$group<- factor(gathered.dna.repair_col.sig_reads$group, levels=c("WT","KO","RE")) ##change levles for plot
```

```{r}
ggboxplot(gathered.dna.repair_col.sig_reads, x = "symbol",
          y = "normalized_counts",
          title="DNA Repair Related Genes",
          color = "group",
          add = "jitter",
          # shape="sampleName",
          # group.by= "cluster",
          merge = TRUE,
          ylab = "Normalized Counts",
          xlab = "",
          legend= "right",
          x.text.angle = 45,
          palette = "aaas")+
  yscale("log10", .format = FALSE)
# graph2ppt(file="figures/DNA_REPAIR.Counts.pptx", width =8,aspectr=sqrt(2),font = "Arial",bg = "transparent",margins =0)
graph2pdf(file="figures/DNA_REPAIR.Counts.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

## 2.6 Plot qPCR results

```{r}
FUS.qPCR.all <- read.csv("/Users/weiyanjia/Desktop/RNA_Seq/FUS_qPCR_RNA_SEQ/FUS_Seq_SSIV_All.csv", header = TRUE)
FUS.qPCR.all.sub <- FUS.qPCR.all %>%
                    dplyr::select(Target, Sample, Expression,Expression.SD)%>%
                    dplyr::filter(!is.na(Expression))%>%
                    dplyr::filter(Target!="WRN",Target!="WNT11",Target!="USP13",Target!="ATF6",Target!="SLC4A10",Target!="POLD4",Target!="SLC26A7",Target!="EPHA6",Target!="SLC6A15")
FUS.qPCR.all.sub$Sample<- factor(FUS.qPCR.all.sub$Sample, levels=c("GUSU2OS","GUSClone110","FUSClone110")) 

FUS.qPCR.DNA.replication<-FUS.qPCR.all.sub%>%
                    dplyr::filter(Target=="GINS4" | Target=="RFC4" | Target=="TIMELESS" | Target=="MCM4" | Target=="RFC3")

FUS.qPCR.DNA.repair<-FUS.qPCR.all.sub%>%
                    dplyr::filter(Target=="53BP1"|Target=="FANCA"|Target=="FANCD2"| Target=="GINS4"|Target=="RFC4")
            
```


```{r fig.height=5, fig.width=10}

q1<-ggplot(FUS.qPCR.all.sub, aes(Target, Expression)) +
  geom_errorbar(
    aes(ymin=Expression-Expression.SD, ymax=Expression+Expression.SD, color = Sample),
    position = position_dodge(0.8), width = 0.8
    )+
  geom_point(aes(color = Sample), position = position_dodge(0.8))


q1+labs(title="qPCR Verification OF FUS RNA-Seq", x="", y = "Realtive expression level")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	theme(plot.title = element_text(hjust = 0.5))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),panel.border = element_blank(), axis.line = element_line(colour = "black"))+
  scale_color_manual(values = c("#3B4992FF","#EE0000FF","#008B45FF"))
# graph2ppt(file="figures/FUS_qPCR.pptx", width =8,aspectr=sqrt(2),font = "Arial",bg = "transparent",margins =0)
graph2pdf(file="figures/FUS_qPCR.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```
```{r}
q2<-ggplot(FUS.qPCR.DNA.replication, aes(Target, Expression)) +
  geom_errorbar(
    aes(ymin=Expression-Expression.SD, ymax=Expression+Expression.SD, color = Sample),
    position = position_dodge(0.8), width = 0.8
    )+
  geom_point(aes(color = Sample), position = position_dodge(0.8))

q2+labs(title="qPCR OF DNA Replication", x="", y = "Realtive expression level")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	theme(plot.title = element_text(hjust = 0.5))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),panel.border = element_blank(), axis.line = element_line(colour = "black"))+
  scale_color_manual(values = c("#3B4992FF","#EE0000FF","#008B45FF"))

# graph2ppt(file="figures/FUS_qPCR_DNA_Replication.pptx", width =8,aspectr=sqrt(2),font = "Arial",bg = "transparent",margins =0)
graph2pdf(file="figures/FUS_qPCR_DNA_Replication.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```


```{r}
q3<-ggplot(FUS.qPCR.DNA.repair, aes(Target, Expression)) +
  geom_errorbar(
    aes(ymin=Expression-Expression.SD, ymax=Expression+Expression.SD, color = Sample),
    position = position_dodge(0.8), width = 0.8
    )+
  geom_point(aes(color = Sample), position = position_dodge(0.8))

q3+labs(title="qPCR OF DNA Repair", x="", y = "Realtive expression level")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	theme(plot.title = element_text(hjust = 0.5))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),panel.border = element_blank(), axis.line = element_line(colour = "black"))+
  scale_color_manual(values = c("#3B4992FF","#EE0000FF","#008B45FF"))

# graph2ppt(file="figures/FUS_qPCR_DNA_Repair.pptx", width =8,aspectr=sqrt(2),font = "Arial",bg = "transparent",margins =0)
graph2pdf(file="figures/FUS_qPCR_DNA_Repair.pdf", width=8, aspectr=sqrt(2),font = "Arial",bg = "transparent")
```

```{r}
sessionInfo()
```

