---
title: "RT Signal Enrichment"
author: "Weiyan"
date: "5/27/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(EnrichedHeatmap) # for making heatmap
library(rtracklayer)  # for reading bigwig and bed files
library(GenomicRanges)
library(circlize)
library(png)
library(dplyr)
library(Hmisc)
library(tidyverse)
```

# Read in bigwig files to GRanges object
```{r}
# read in bigwig files to GRanges object, it can be slow. bigwig is several hundred MB
# you can use which argument to restrict the data in certain regions.
U2OS.bg<- import("data/U2OS_RT_R1-X_Loess_smoothing.bedGraph", format = "BedGraph")
Clone110.bg<- import("data/Clone110_RT_R1-X_Loess_smoothing.bedGraph", format = "BedGraph")
FUSClone110.bg<- import("data/FUSClone110_RT_R1-X_Loess_smoothing.bedGraph", format = "BedGraph")
```

# 1. ERD
## 1.1 load data
```{r}
# read in the bed peak files to GRanges object
ERD.bed<- import("data/ERD_lost.bed", format = "BED")
```

## 1.2 take only the center of the peaks

```{r}
ERD.1MB<- resize(ERD.bed, width = 1000000, fix = "center")
ERD.1MB.center<- resize(ERD.1MB, width =1, fix = "center")
```

## 1.3 prepare matrix
```{r}
U2OS.mat<- normalizeToMatrix(U2OS.bg, ERD.1MB.center, value_column = "score",
                             mean_mode="w0", w=1000, extend = 500000)

Clone110.mat<- normalizeToMatrix(Clone110.bg, ERD.1MB.center, value_column = "score",
                                 mean_mode="w0", w=1000, extend = 500000)

FUSClone110.mat<- normalizeToMatrix(FUSClone110.bg, ERD.1MB.center, value_column = "score",
                                    mean_mode="w0", w=1000, extend = 500000)

```

### 1.3.1 check data range for "keep" argument in normalizeToMatrix
```{r}
quantile(U2OS.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(Clone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(FUSClone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
```

## 1.4 mapping colors and enrich plot
**from the quantile, I choose the color mapping range**
```{r fig.height=8, fig.width=6}
col_fun_U2OS<- circlize::colorRamp2(c(0, 1), c("white", "red"))
col_fun_Clone110<- circlize::colorRamp2(c(0, 1), c("white", "red"))
col_fun_FUSClone110<- circlize::colorRamp2(c(0, 1), c("white", "red"))

 EnrichedHeatmap(U2OS.mat, axis_name_rot = 0, name = "U2OS",
                column_title = "U2OS", use_raster = TRUE, col = col_fun_U2OS,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
     side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(Clone110.mat, axis_name_rot = 0, name = "Clone110", 
                  column_title = "Clone110", use_raster = TRUE, col = col_fun_Clone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(FUSClone110.mat, axis_name_rot = 0, name = "FUSClone110", 
                  column_title = "FUSClone110", use_raster = TRUE, col = col_fun_FUSClone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))
```

## 1.5 meta profile plot
```{r}
U2OS_mean<- data.frame(avg = colMeans(U2OS.mat), 
                       CI_lower = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[2,],
                       CI_upper = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[3,]) %>% 
  mutate(factor = "U2OS", pos = colnames(U2OS.mat)) 

Clone110_mean<- data.frame(avg = colMeans(Clone110.mat), 
                           CI_lower = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                           CI_upper = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "Clone110", pos = colnames(Clone110.mat))

FUSClone110_mean<- data.frame(avg = colMeans(FUSClone110.mat), 
                              CI_lower = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                              CI_upper = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "FUSClone110", pos = colnames(FUSClone110.mat))

combine_all<- bind_rows(U2OS_mean, Clone110_mean, FUSClone110_mean)

## change position to factor and order it 
combine_all$pos<- factor(combine_all$pos, levels= U2OS_mean$pos)
```

### 1.5.1 plot meta profile (without confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("ERD") + 
  ylab("RT")+
  ggtitle("RT signal in ERD")
```

### 1.5.2 plot meta profile (with confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  geom_ribbon(aes(ymin= CI_lower, ymax=CI_upper), alpha=0.2, col = "#8B7E66") +
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("ERD") + 
  ylab("RT")+
  ggtitle("RT signal in ERD")
```

# 2. MRD
## 2.1 load data
```{r}
# read in the bed peak files to GRanges object
MRD.bed<- import("data/MRD_lost.bed", format = "BED")
```

## 2.2 take only the center of the peaks

```{r}
MRD.1MB<- resize(MRD.bed, width = 1000000, fix = "center")
MRD.1MB.center<- resize(MRD.1MB, width =1, fix = "center")
```

## 2.3 prepare matrix
```{r}
U2OS.mat<- normalizeToMatrix(U2OS.bg, MRD.1MB.center, value_column = "score",
                             mean_mode="w0", w=1000, extend = 500000)

Clone110.mat<- normalizeToMatrix(Clone110.bg, MRD.1MB.center, value_column = "score",
                                 mean_mode="w0", w=1000, extend = 500000)

FUSClone110.mat<- normalizeToMatrix(FUSClone110.bg, MRD.1MB.center, value_column = "score",
                                    mean_mode="w0", w=1000, extend = 500000)

```

### 2.3.1 check data range for "keep" argument in normalizeToMatrix
```{r}
quantile(U2OS.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(Clone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(FUSClone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
```

## 2.4 mapping colors and enrich plot
**from the quantile, I choose the color mapping range**
```{r fig.height=8, fig.width=6}
col_fun_U2OS<- circlize::colorRamp2(c(-0.4, 0.4), c("white", "red"))
col_fun_Clone110<- circlize::colorRamp2(c(-0.4, 0.4), c("white", "red"))
col_fun_FUSClone110<- circlize::colorRamp2(c(-0.4, 0.4), c("white", "red"))

 EnrichedHeatmap(U2OS.mat, axis_name_rot = 0, name = "U2OS",
                column_title = "U2OS", use_raster = TRUE, col = col_fun_U2OS,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
     side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(Clone110.mat, axis_name_rot = 0, name = "Clone110", 
                  column_title = "Clone110", use_raster = TRUE, col = col_fun_Clone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(FUSClone110.mat, axis_name_rot = 0, name = "FUSClone110", 
                  column_title = "FUSClone110", use_raster = TRUE, col = col_fun_FUSClone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))
```

## 2.5 meta profile plot
```{r}
U2OS_mean<- data.frame(avg = colMeans(U2OS.mat), 
                       CI_lower = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[2,],
                       CI_upper = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[3,]) %>% 
  mutate(factor = "U2OS", pos = colnames(U2OS.mat)) 

Clone110_mean<- data.frame(avg = colMeans(Clone110.mat), 
                           CI_lower = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                           CI_upper = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "Clone110", pos = colnames(Clone110.mat))

FUSClone110_mean<- data.frame(avg = colMeans(FUSClone110.mat), 
                              CI_lower = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                              CI_upper = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "FUSClone110", pos = colnames(FUSClone110.mat))

combine_all<- bind_rows(U2OS_mean, Clone110_mean, FUSClone110_mean)

## change position to factor and order it 
combine_all$pos<- factor(combine_all$pos, levels= U2OS_mean$pos)
```

### 2.5.1 plot meta profile (without confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("MRD") + 
  ylab("RT")+
  ggtitle("RT signal in MRD")
```

### 2.5.2 plot meta profile (with confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  geom_ribbon(aes(ymin= CI_lower, ymax=CI_upper), alpha=0.2, col = "#8B7E66") +
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("MRD") + 
  ylab("RT")+
  ggtitle("RT signal in MRD")
```
# 3. LRD
## 3.1 load data
```{r}
# read in the bed peak files to GRanges object
LRD.bed<- import("data/LRD_lost.bed", format = "BED")
```

## 3.2 take only the center of the peaks

```{r}
LRD.1MB<- resize(LRD.bed, width = 1000000, fix = "center")
LRD.1MB.center<- resize(LRD.1MB, width =1, fix = "center")
```

## 3.3 prepare matrix
```{r}
U2OS.mat<- normalizeToMatrix(U2OS.bg, LRD.1MB.center, value_column = "score",
                             mean_mode="w0", w=1000, extend = 500000)

Clone110.mat<- normalizeToMatrix(Clone110.bg, LRD.1MB.center, value_column = "score",
                                 mean_mode="w0", w=1000, extend = 500000)

FUSClone110.mat<- normalizeToMatrix(FUSClone110.bg, LRD.1MB.center, value_column = "score",
                                    mean_mode="w0", w=1000, extend = 500000)

```

### 3.3.1 check data range for "keep" argument in normalizeToMatrix
```{r}
quantile(U2OS.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(Clone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
quantile(FUSClone110.mat, probs = c(0, 0.25, 0.5, 0.75, 0.99, 1))
```

## 3.4 mapping colors and enrich plot
**from the quantile, I choose the color mapping range**
```{r fig.height=8, fig.width=6}
col_fun_U2OS<- circlize::colorRamp2(c(-1, 0), c("white", "red"))
col_fun_Clone110<- circlize::colorRamp2(c(-1, 0), c("white", "red"))
col_fun_FUSClone110<- circlize::colorRamp2(c(-1, 0), c("white", "red"))

 EnrichedHeatmap(U2OS.mat, axis_name_rot = 0, name = "U2OS",
                column_title = "U2OS", use_raster = TRUE, col = col_fun_U2OS,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
     side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(Clone110.mat, axis_name_rot = 0, name = "Clone110", 
                  column_title = "Clone110", use_raster = TRUE, col = col_fun_Clone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))+
  
  EnrichedHeatmap(FUSClone110.mat, axis_name_rot = 0, name = "FUSClone110", 
                  column_title = "FUSClone110", use_raster = TRUE, col = col_fun_FUSClone110,
                  top_annotation = HeatmapAnnotation(lines = anno_enriched(axis_param = list(facing ="inside",
       side ="left"                                                                                       ))))
```

## 3.5 meta profile plot
```{r}
U2OS_mean<- data.frame(avg = colMeans(U2OS.mat), 
                       CI_lower = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[2,],
                       CI_upper = apply(U2OS.mat, 2, Hmisc::smean.cl.normal)[3,]) %>% 
  mutate(factor = "U2OS", pos = colnames(U2OS.mat)) 

Clone110_mean<- data.frame(avg = colMeans(Clone110.mat), 
                           CI_lower = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                           CI_upper = apply(Clone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "Clone110", pos = colnames(Clone110.mat))

FUSClone110_mean<- data.frame(avg = colMeans(FUSClone110.mat), 
                              CI_lower = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[2,],
                              CI_upper = apply(FUSClone110.mat, 2, Hmisc::smean.cl.normal)[3,]) %>%
  mutate(factor = "FUSClone110", pos = colnames(FUSClone110.mat))

combine_all<- bind_rows(U2OS_mean, Clone110_mean, FUSClone110_mean)

## change position to factor and order it 
combine_all$pos<- factor(combine_all$pos, levels= U2OS_mean$pos)
```

### 3.5.1 plot meta profile (without confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("LRD") + 
  ylab("RT")+
  ggtitle("RT signal in LRD")
```

### 3.5.2 plot meta profile (without confidence interval)
```{r}
ggplot(combine_all, aes(x = pos,y = avg, group = factor)) + geom_line(aes(color = factor)) + 
  geom_ribbon(aes(ymin= CI_lower, ymax=CI_upper), alpha=0.2, col = "#8B7E66") +
  theme_bw(base_size = 14) +
  theme(axis.ticks.x = element_blank()) +
  scale_x_discrete(breaks = c("u1", "d10", "d500"), labels =c ("-500kb", "center", "500kb")) +
  xlab("LRD") + 
  ylab("RT")+
  ggtitle("RT signal in LRD")
```

```{r}
sessionInfo()
```

